---
title: "Introduction to SNSeg and Examples"
output: rmarkdown::html_vignette
description: >
  Start here if this is your first time using SNSeg. You'll learn the basic   
  philosophy, the type of parameters used for change-points detection, and 
  examples for functions within this R package.
vignette: >
  %\VignetteIndexEntry{SNSeg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(SNSeg)
```

SNSeg supports change-points estimation for both univariate and multivariate time series (including high-dimensional time series with dimension greater than 10.) using Self-Normalization (SN) based framework. Please read Zhao, Jiang and Shao (2022) <doi.org/10.1111/rssb.12552> for details of the SN-based algorithms.

The package contain three functions for change-points estimation:

* The function `SNSeg_Uni` works for a univariate time series.
* The function `SNSeg_Multi` works for multivariate time series with dimension up to ten.
* The function `SNSeg_HD` works for high-dimensional time series with dimension above ten.

All functions contain two important input arguments: `grid_size_scale` and `grid_size`.

* `grid_size_scale`: the trimming parameter $\epsilon$, a parameter to control the local window size `grid_size`. The range of `grid_size_scale` should be between 0.05 and 0.5. Any input `grid_size_scale` smaller than 0.05 will be automatically changed to 0.05, and any input `grid_size_scale` greater than 0.5 will be automatically changed to 0.5.
* `grid_size`: the local window size $h$ for local scanning for each time point. According to Zhao et al. (2022), `grid_size` is a product of `grid_size_scale` and the length of time.

Users can set their own `grid_size` or leave it as `NULL` in the input arguments. If `grid_size` is `NULL`, these functions will compute the SN test statistic using the value of `grid_size_scale`. If `grid_size` is set by users, the functions will first calculate `grid_size_scale` by diving the length of time, and then compute the SN test statistic.

For the other input arguments:
* `ts`: Users should enter a time series for this argument. For `SNSeg_Uni`, the dimension of `ts` should be exactly one in most cases (or two when `paras_to_test = 'bivcor'`); for `SNSeg_Multi`, the dimension must be at least two but no more than ten; for `SNSeg_HD`, the dimension must be greater than ten.
* `paras_to_test`: This argument in functions `SNSeg_Uni` and `SNSeg_Multi` allows users to enter the parameter(s) they would like to test the change in.
* `confidence`: Users should choose a confidence level among 0.9, 0.95, 0.99, 0.995 and 0.995. A smaller confidence level is easier to reject the null hypothesis and detect a change-point. 

The package also offers an option to plot the time series (this only works for the univariate time series cases!) Users need to set `plot_SN = TRUE` to visualize the time series, and `est_cp_loc = TRUE` to add the estimated change-point locations in the plot.

## SN Test Statistic Plot: `max_SNsweep`
To visualize the computed SN-based test statistic at each time point, users can apply the function `max_SNsweep` by plugging in the output object from one of the functions `SNSeg_Uni`, `SNSeg_Multi` and `SNSeg_HD`. The options `est_cp_loc = TRUE` and `critical_loc = TRUE` are provided to draw the estimated change-point locations and the critical value threshold inside the test statistic plot. 

`max_SNsweep` also returns the SN-based test statistic for each time point. A large number of test statistics in the output can be messy for users who only seek to generate a plot. To hide the test statistics output, users can create an arbitrary variable name and set it to the `max_SNsweep` function, e.g., `SN_stat <- max_SNsweep(...)`. 

## Parameter Estimates of Each Segment Separated by the Detected Change-Points: `SNSeg_estimate`
The function `SNSeg_estimate` allows users to compute the parameter estimates (e.g., mean, variance, acf, quantile, etc.) of each of the segments separated by the estimated change-points. To use this function, users should use the output of the functions `SNSeg_Uni`, `SNSeg_Multi` and `SNSeg_HD` as the input of `SNSeg_estimate`.

## S3 methods: summary, print and plot

The typical S3 methods `summary`, `print` and `plot` are available to `SNSeg_Uni`, `SNSeg_Multi` and `SNSeg_HD` objects. The `summary` method displays the parameter to be tested, the estimated change-point amount and locations, the `grid_size`, confidence level as well as the critical value of the SN-based test. The `print` method shows the change-point locations. The `plot` method plots the time series, and similar to the argument `plot_SN = TRUE`, the `plot` method allows users to generate time series segmentation plot(s) with the estimated change-point locations. It also provides `n_plot` option to allow users to generate any number of plots they want. Users can apply their preferred color of the change-point(s) within the plot(s) by setting `cpts.col` to any color.

We then provide the examples of the functions `SNSeg_Uni`, `SNSeg_Multi` and `SNSeg_HD`. 

## Examples of `SNSeg_Uni`: 

The function `SNSeg_Uni` detect change-points for a univariate time series based on the change in a single or parameters. 

* For a univariate time series, the input argument `paras_to_test` allows one or a combination of multiple parameters from "mean", "variance", "acf" and a numeric quantile value between 0 and 1. For instance, to test the change in autocorrelation and 80th quantile, users should set `paras_to_test` to c('acf', 0.8). 
* To test the change in correlation between two univariate time series, the input time series must be two-dimensional and `paras_to_test` should be set to `bivcor`.
* Users can also set `paras_to_test` using their own parameters. In this case, the input of `paras_to_test` needs to be a function which returns a numeric value.  

We provide examples for different cases:

```{r}
# Please run the following function before running examples:
mix_GauGPD <- function(u,p,trunc_r,gpd_scale,gpd_shape){
  indicator <- u<p
  rv <- rep(0, length(u))
  rv[indicator>0] <- qtruncnorm(u[indicator>0]/p,a=-Inf,b=trunc_r)
  rv[indicator<=0] <- qgpd((u[indicator<=0]-p)/(1-p), loc=trunc_r, scale=gpd_scale,shape=gpd_shape)
  return(rv)
}
```

## Test in a single parameter
### Segmentation for Mean
```{r}
set.seed(7)
n <- 2000
reptime <- 2
cp_sets <- round(n*c(0,cumsum(c(0.5,0.25)),1))
mean_shift <- c(0.4,0,0.4)
rho <- -0.7
ts <- MAR(n, reptime, rho)
no_seg <- length(cp_sets)-1
for(index in 1:no_seg){ # Mean shift
  tau1 <- cp_sets[index]+1   
  tau2 <- cp_sets[index+1]
  ts[tau1:tau2,] <- ts[tau1:tau2,] + mean_shift[index]
}
ts <- ts[,2]
# grid_size undefined
result <- SNSeg_Uni(ts, paras_to_test = "mean", confidence = 0.9,
                    grid_size_scale = 0.05, grid_size = NULL, 
                    plot_SN = FALSE, est_cp_loc = FALSE)
# grid_size defined & generate time series segmentation plot
result <- SNSeg_Uni(ts, paras_to_test = "mean", confidence = 0.9,
                    grid_size_scale = 0.05, grid_size = 116, 
                    plot_SN = TRUE, est_cp_loc = TRUE)
# Estimated change-point locations
result$est_cp
# Parameter estimates (mean) of each segment
SNSeg_estimate(result)
# plot the SN-based test statistic
SN_stat <- max_SNsweep(result, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

We then show how to use the S3 methods `summary`, `print` and `plot`.

```{r}
summary(result)
```

```{r}
print(result)
```

```{r}
plot(result, cpts.col = 'red')
```

We can see both the `plot_SN = TRUE` option and the `plot.SN` function generates the same plot. Users can select any choice based on their preferences. The class of the argument `result` is `SNSeg_Uni`.

### Segmentation for Variance
```{r}
set.seed(7)
ts <- MAR_Variance(2, "V1")
ts <- ts[,2]
# grid_size defined
result <- SNSeg_Uni(ts, paras_to_test = "variance", confidence = 0.9,
                    grid_size_scale = 0.05, grid_size = NULL, 
                    plot_SN = FALSE, est_cp_loc = TRUE)
# Estimated change-point locations
result$est_cp
# Parameter estimates (variance) of each segment
SNSeg_estimate(result)
# plot the SN-based test statistic
SN_stat <- max_SNsweep(result, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result)
print(result)
plot(result, cpts.col = 'red')
```

### Segmentation for Autocorrelation
```{r}
set.seed(7)
ts <- MAR_Variance(2, "A3")
ts <- ts[,2]
# grid_size defined
result <- SNSeg_Uni(ts, paras_to_test = "acf", confidence = 0.9,
          grid_size_scale = 0.05, grid_size = 92, plot_SN = FALSE,
          est_cp_loc = TRUE)
# Estimated change-point locations
result$est_cp
# Parameter estimates (acf) of each segment
SNSeg_estimate(result)
# plot the SN-based test statistic
SN_stat <- max_SNsweep(result, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result)
print(result)
plot(result, cpts.col = 'red')
```

### Segmentation for bivariate correlation
```{r}
library(mvtnorm)
set.seed(7)
n <- 1000
sigma_cross <- list(4*matrix(c(1,0.8,0.8,1), nrow=2),
                    matrix(c(1,0.2,0.2,1), nrow=2),
                    matrix(c(1,0.8,0.8,1), nrow=2))
cp_sets <- round(c(0,n/3,2*n/3,n))
noCP <- length(cp_sets)-2
rho_sets <- rep(0.5, noCP+1)
ts <- MAR_MTS_Covariance(n, 2, rho_sets, cp_sets, sigma_cross)
ts <- ts[1][[1]]
# grid_size defined
result <- SNSeg_Uni(ts, paras_to_test = "bivcor", confidence = 0.9,
                    grid_size_scale = 0.05, grid_size = 77, 
                    plot_SN = FALSE, est_cp_loc = TRUE)
# Estimated change-point locations
result$est_cp
# Parameter estimates (bivariate correlation) of each segment
SNSeg_estimate(result)
# plot the SN-based test statistic
SN_stat <- max_SNsweep(result, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result)
print(result)
plot(result, cpts.col = 'red')
```

### Segmentation for quantile
```{r}
library(truncnorm)
library(evd)
set.seed(7)
n <- 1000
cp_sets <- c(0,n/2,n)
noCP <- length(cp_sets)-2
reptime <- 2
rho <- 0.2
# AR time series with no change-point (mean, var)=(0,1)
ts <- MAR(n, reptime, rho)*sqrt(1-rho^2)
trunc_r <- 0
p <- pnorm(trunc_r)
gpd_scale <- 2
gpd_shape <- 0.125
for(ts_index in 1:reptime){
  ts[(cp_sets[2]+1):n, ts_index] <- mix_GauGPD(pnorm(ts[(cp_sets[2]+1):n, ts_index]),
p,trunc_r,gpd_scale,gpd_shape)
}
ts <- ts[,2]
# grid_size undefined
# test in 90% quantile
result <- SNSeg_Uni(ts, paras_to_test = 0.9, confidence = 0.9,
                    grid_size_scale = 0.066, grid_size = NULL,
                    plot_SN = FALSE, est_cp_loc = TRUE)
# Estimated change-point locations
result$est_cp
# Parameter estimates (90th quantile) of each segment
SNSeg_estimate(result)
# plot the SN-based test statistic
SN_stat <- max_SNsweep(result, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result)
print(result)
plot(result, cpts.col = 'red')
```

### Test in a general functional
```{r}
set.seed(7)
n <- 500
reptime <- 2
cp_sets <- round(n*c(0,cumsum(c(0.5,0.25)),1))
mean_shift <- c(0.4,0,0.4)
rho <- -0.7
ts <- MAR(n, reptime, rho)
no_seg <- length(cp_sets)-1
for(index in 1:no_seg){ # Mean shift
  tau1 <- cp_sets[index]+1
  tau2 <- cp_sets[index+1]
  ts[tau1:tau2,] <- ts[tau1:tau2,] + mean_shift[index]
}
ts <- ts[,2]
# set a general functional for the input 'paras_to_test'
paras_to_test = function(ts){
  mean(ts)
}
result.SNCP.general <- SNSeg_Uni(ts, paras_to_test = paras_to_test, 
                                 confidence = 0.9, grid_size_scale = 0.05, 
                                 grid_size = NULL, plot_SN = FALSE, 
                                 est_cp_loc = TRUE)
# Estimated change-point locations
result.SNCP.general$est_cp
# Parameter estimates (general functional) of each segment
SNSeg_estimate(result.SNCP.general)
# plot the SN-based test statistic
SN_stat <- max_SNsweep(result.SNCP.general, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result.SNCP.general)
print(result.SNCP.general)
plot(result.SNCP.general, cpts.col = 'red')
```

### Test in multiple parameters
```{r}
set.seed(7)
n <- 1000
cp_sets <- c(0,333,667,1000)
no_seg <- length(cp_sets)-1
rho <- 0
# AR time series with no change-point (mean, var)=(0,1)
ts <- MAR(n, 2, rho)*sqrt(1-rho^2)
no_seg <- length(cp_sets)-1
sd_shift <- c(1,1.6,1)
for(index in 1:no_seg){ # Mean shift
  tau1 <- cp_sets[index]+1
  tau2 <- cp_sets[index+1]
  ts[tau1:tau2,] <- ts[tau1:tau2,]*sd_shift[index]
}
d <- 2
ts <- ts[,2]

# Test in 90th and 95th quantile with grid_size undefined
result <- SNSeg_Uni(ts, paras_to_test = c(0.9, 0.95), confidence = 0.9,
                    grid_size_scale = 0.05, grid_size = NULL, 
                    plot_SN = FALSE, est_cp_loc = FALSE)

# Test in 90th quantile and the variance with grid_size undefined
result <- SNSeg_Uni(ts, paras_to_test = c(0.9, 'variance'),
                    confidence = 0.95, grid_size_scale = 0.078,
                    grid_size = NULL, plot_SN = FALSE, 
                    est_cp_loc = FALSE)

# Test in 90th quantile, variance and acf with grid_size undefined
result <- SNSeg_Uni(ts, paras_to_test = c(0.9,'variance', "acf"),
                    confidence = 0.9, grid_size_scale = 0.064,
                    grid_size = NULL, plot_SN = TRUE, 
                    est_cp_loc = TRUE)
# Estimated change-point locations
result$est_cp
```

```{r}
# Test in 60th quantile, mean, variance and acf with grid_size defined
result.last <- SNSeg_Uni(ts, paras_to_test = c(0.6, 'mean', "variance",   
                    "acf"), confidence = 0.9, grid_size_scale = 0.05,
                    grid_size = 83, plot_SN = FALSE, est_cp_loc = TRUE)
# Estimated change-point locations
result.last$est_cp
# Parameter estimates (of the last example) of each segment
SNSeg_estimate(result.last)
# SN-based test statistic segmentation plot
SN_stat <- max_SNsweep(result.last, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result.last)
print(result.last)
plot(result.last, cpts.col = 'red')
```

## Examples: `SNSeg_Multi`
The function `SNSeg_Multi` detects change-points for multivariate time series based on the change in multivariate means or covariances. Users can set `paras_to_test` to `mean` or `covariance` for change-points estimation. 

* The dimension of parameters to be tested must be at most ten. If the parameter is `mean`, the dimension is equivalent to the number of time series. If the parameter is `covariance`, the dimension is equivalent to the number of unknown parameters within the covariance matrix.

For examples of multivariate means and covariances, please check the following commands:

```{r}
# Please run this function before running examples:
exchange_cor_matrix <- function(d, rho){
  tmp <- matrix(rho, d, d)
  diag(tmp) <- 1
  return(tmp)
}
```

### Segmentation for Multivariate Mean
```{r}
library(mvtnorm)
set.seed(10)
d <- 5
n <- 1000
cp_sets <- round(n*c(0,cumsum(c(0.075,0.3,0.05,0.1,0.05)),1))
mean_shift <- c(-3,0,3,0,-3,0)/sqrt(d)
mean_shift <- sign(mean_shift)*ceiling(abs(mean_shift)*10)/10
rho_sets <- 0.5
sigma_cross <- list(exchange_cor_matrix(d,0))
ts <- MAR_MTS_Covariance(n, 2, rho_sets, cp_sets=c(0,n), sigma_cross)
noCP <- length(cp_sets)-2
no_seg <- length(cp_sets)-1
for(rep_index in 1:2){
  for(index in 1:no_seg){ # Mean shift
    tau1 <- cp_sets[index]+1
    tau2 <- cp_sets[index+1]
    ts[[rep_index]][,tau1:tau2] <- ts[[rep_index]][,tau1:tau2] + mean_shift[index]
    }
}
ts <- ts[1][[1]]

# grid_size undefined
result <- SNSeg_Multi(ts, paras_to_test = "mean", confidence = 0.95,
                      grid_size_scale = 0.079, grid_size = NULL,
                      plot_SN = FALSE, est_cp_loc = TRUE)
# grid_size defined
result <- SNSeg_Multi(ts, paras_to_test = "mean", confidence = 0.99,
                      grid_size_scale = 0.05, grid_size = 65,
                      plot_SN = FALSE, est_cp_loc = TRUE)
# Estimated change-point locations
result$est_cp
# Parameter estimates (multivariate mean) of each segment
SNSeg_estimate(result)
# SN-based test statistic segmentation plot
SN_stat <- max_SNsweep(result, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result)
print(result)
par(mfrow=c(2,3))
plot(result, cpts.col = 'red')
```

### Segmentation for Multivariate Covariance
```{r}
library(mvtnorm)
set.seed(10)
reptime <- 2
d <- 4
n <- 1000
sigma_cross <- list(exchange_cor_matrix(d,0.2),
                    2*exchange_cor_matrix(d,0.5),
                    4*exchange_cor_matrix(d,0.5))
rho_sets <- c(0.3,0.3,0.3)
mean_shift <- c(0,0,0) # with mean change
cp_sets <- round(c(0,n/3,2*n/3,n))
ts <- MAR_MTS_Covariance(n, reptime, rho_sets, cp_sets, sigma_cross)
noCP <- length(cp_sets)-2
no_seg <- length(cp_sets)-1
for(rep_index in 1:reptime){
  for(index in 1:no_seg){ # Mean shift
    tau1 <- cp_sets[index]+1
    tau2 <- cp_sets[index+1]
    ts[[rep_index]][,tau1:tau2] <- ts[[rep_index]][,tau1:tau2] +
      mean_shift[index]
    }
}
ts <- ts[[1]]

# grid_size undefined
result <- SNSeg_Multi(ts, paras_to_test = "covariance", 
                      confidence = 0.9, grid_size_scale = 0.05, 
                      grid_size = NULL, plot_SN = FALSE, est_cp_loc = FALSE)
# grid_size defined
result <- SNSeg_Multi(ts, paras_to_test = "covariance", 
                      confidence = 0.9, grid_size_scale = 0.05, 
                      grid_size = 81, plot_SN = FALSE,
                      est_cp_loc = TRUE)
# Estimated change-point locations
result$est_cp
# Parameter estimates (covariance estimate) of each segment
SNSeg_estimate(result)
# SN-based test statistic segmentation plot
SN_stat <- max_SNsweep(result, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result)
print(result)
par(mfrow=c(1,1))
plot(result, cpts.col = 'red')
```

## Examples: `SNSeg_HD`
The function `SNSeg_HD` performs change-points estimation for high-dimensional time series (dimension greater than 10) using the change in high-dimensional means. For usage examples of `SNSeg_HD`, we provide the followig code:

```{r}
n <- 600
p <- 100
nocp <- 5
cp_sets <- round(seq(0,nocp+1,1)/(nocp+1)*n)
num_entry <- 5
kappa <- sqrt(4/5) # Wang et al(2020)
mean_shift <- rep(c(0,kappa),100)[1:(length(cp_sets)-1)]
set.seed(1)
ts <- matrix(rnorm(n*p,0,1),n,p)
no_seg <- length(cp_sets)-1
for(index in 1:no_seg){ # Mean shift
  tau1 <- cp_sets[index]+1
  tau2 <- cp_sets[index+1]
  ts[tau1:tau2,1:num_entry] <- ts[tau1:tau2,1:num_entry] +
    mean_shift[index] # sparse change
}
# SN segmentation plot (plot the first 3 time series)
# grid_size undefined
result <- SNSeg_HD(ts, confidence = 0.9, grid_size_scale = 0.05,
                   grid_size = NULL, plot_SN = FALSE, est_cp_loc = TRUE,
                   n_plot = 3)
# grid_size defined
result <- SNSeg_HD(ts, confidence = 0.9, grid_size_scale  = 0.05,
                   grid_size = 52, plot_SN = FALSE, est_cp_loc = TRUE,
                   n_plot = 3)
# Estimated change-point locations
result$est_cp
# Parameter estimates (high-dimensional means) of each segment
summary.stat <- SNSeg_estimate(result)
# SN-based test statistic segmentation plot
SN_stat <- max_SNsweep(result, plot_SN = TRUE, est_cp_loc = TRUE,
                       critical_loc = TRUE)
```

```{r}
summary(result)
print(result)
plot(result, cpts.col = 'red')
```
We note that only the first three time series were plotted by setting `n_plot = 3`.
